{\em Cloudbreak}: Accurate and Scalable Genomic Structural Variation Detection in the Cloud with MapReduce

### 2013-07-10 manuscript/arxiv.tex
final arxiv version

The detection of genomic structural variations (SV) remains a difficult challenge in analyzing sequencing data, and the growing size and number of sequenced genomes have rendered SV detection a bona fide big data problem. MapReduce is a proven, scalable solution for distributed computing on huge data sets. We describe a conceptual framework for SV detection algorithms in MapReduce based on computing local genomic features, and use it to develop a deletion and insertion detection algorithm, Cloudbreak. On simulated and real data sets, Cloudbreak achieves accuracy improvements over popular SV detection algorithms, and accurately genotypes variants from diploid samples. It provides dramatically shorter runtimes and the ability to scale to big data volumes on large compute clusters. Cloudbreak includes tools to set up and configure MapReduce (Hadoop) clusters on cloud services, enabling on-demand cluster computing. Our implementation and source code are available at \url{http://github.com/cwhelan/cloudbreak}.

@@

\setcounter{table}{1}
\setcounter{figure}{0}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}

@@

This should compile the code, execute tests, and create a distribution file,
 \texttt{cloudbreak-\$VERSION-dist.tar.gz} in the \texttt{target\slash } directory. You can then copy
 that distribution file to somewhere else on your system, unpack it with
 \texttt{tar -xzvf cloudbreak-\$VERSION-dist.tar.gz} and access the Cloudbreak jar file
 and related scripts and properties files.

>>>>

The detection of genomic structural variations (SV) remains a difficult challenge in analyzing sequencing data, and the growing size and number of sequenced genomes have rendered SV detection a bona fide big data problem. MapReduce is a proven, scalable solution for distributed computing on huge data sets. We describe a conceptual framework for SV detection algorithms in MapReduce based on computing local genomic features, and use it to develop a deletion and insertion detection algorithm, Cloudbreak. On simulated and real data sets, Cloudbreak achieves accuracy improvements over popular SV detection algorithms, and genotypes variants from diploid samples. It provides dramatically shorter runtimes and the ability to scale to big data volumes on large compute clusters. Cloudbreak includes tools to set up and configure MapReduce (Hadoop) clusters on cloud services, enabling on-demand cluster computing. Our implementation and source code are available at \url{http://github.com/cwhelan/cloudbreak}.

@@

\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}

@@

This should compile the code, execute tests, and create a distribution file, \linebreak
 \texttt{cloudbreak-\$VERSION-dist.tar.gz}, in the \texttt{target\slash } directory. You can then copy
 that distribution file to somewhere else on your system, unpack it with:

 \texttt{tar -xzvf cloudbreak-\$VERSION-dist.tar.gz}

and access the Cloudbreak jar file and related scripts and properties files.
